---
title: "Reading and writing xml files in R"
author: "Pep Porrà"
date: "2021-05-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

Explain pros ans cons of **XML** package and **xml2** package to write and read xml files

## Example

```{r}
library(xml2)
library(XML)
library(tidyverse)
```

## **xml2** package

```{r}
doc1 <- xml2::read_xml('<gpx version="1.1" creator="R">
<trk>
<trkseg>
<trkpt lat="40.779" lon="-74.428" />
<trkpt lat="40.777" lon="-74.418" />
</trkseg>
 </trk>
 <trk>
 <trkseg>
<trkpt lat="50.779" lon="-64.428" />
<trkpt lat="50.777" lon="-64.418" />
</trkseg>
</trk>
</gpx>')
```

Function `xml_children` returns a nodeset. Each node can be accessed as a list element

```{r}
tracks_nodes <- xml2::xml_children(doc1)
tracks_nodes
```

First node within track_nodes

```{r}
tracks_nodes %>% .[[1]]
```

```{r}
xml2::xml_length(doc1)
```
```{r}
xml2::xml_name(doc1)
```
Extract attributes lat and lon

```{r}
xml_attrs(xml_find_all(doc1, "//trkpt"))
```

## Example

```{r}
df <- structure(list(lon = c(-79.045899, -79.045919, -79.045937, -79.045951, 
-79.045967, -79.046174, -79.04619, -79.046203, -79.046302, -79.046311, 
-79.046704, -79.046694, -79.046687, -79.046702, -79.046727, -79.046735, 
-79.046739, -79.046752, -79.046879, -79.046885), lat = c(35.898049, 
35.89805, 35.898054, 35.898059, 35.898066, 35.8981, 35.898108, 
35.898115, 35.898169, 35.898177, 35.898017, 35.898038, 35.898021, 
35.89801, 35.898004, 35.897989, 35.897964, 35.897954, 35.897897, 
35.897905), ele = c("99.6", "99.6", "99.8", "99.8", "99.8", "101.2", 
"101.2", "101.2", "101.6", "102.0", "105.8", "134.2", "134.2", 
"134.2", "107.2", "107.0", "107.2", "107.4", "107.6", "107.6"
), time = c("2019-02-06T01:34:35Z", "2019-02-06T01:34:36Z", "2019-02-06T01:34:37Z", 
"2019-02-06T01:34:38Z", "2019-02-06T01:34:39Z", "2019-02-06T01:34:52Z", 
"2019-02-06T01:34:53Z", "2019-02-06T01:34:54Z", "2019-02-06T01:35:02Z", 
"2019-02-06T01:35:07Z", "2019-02-06T00:15:59Z", "2019-02-06T00:16:00Z", 
"2019-02-06T00:16:01Z", "2019-02-06T00:16:03Z", "2019-02-06T00:16:04Z", 
"2019-02-06T00:16:05Z", "2019-02-06T00:16:09Z", "2019-02-06T00:16:10Z", 
"2019-02-06T00:16:15Z", "2019-02-06T00:16:17Z"), activity = c("Run", 
"Run", "Run", "Run", "Run", "Run", "Run", "Run", "Run", "Run", 
"Run", "Run", "Run", "Run", "Run", "Run", "Run", "Run", "Run", 
"Run")), row.names = c(1020L, 1021L, 1022L, 1023L, 1024L, 1025L, 
1026L, 1027L, 1028L, 1029L, 1L, 2L, 3L, 4L, 5L, 6L, 7L, 8L, 9L, 
10L), class = "data.frame")
```

```{r}
doc <- xml_new_root("gpx",
  version = "1.1.0",
  creator = "R pep",
  "xmlns:xsi"="http://www.w3.org/2001/XMLSchema-instance",
  xmlns="http://www.topografix.com/GPX/1/1",
  "xsi:schemaLocation"="http://www.topografix.com/GPX/1/1 http://www.topografix.com/GPX/1/1/gpx.xsd",
  "xmlns:gpxtpx"="http://www.garmin.com/xmlschemas/TrackPointExtension/v1"
)
```

```{r}
doc
```

```{r}
xml_add_child(doc, "trk")
xml_add_child(doc, "trk")
```

```{r}
doc
```

```{r}
xml_children(doc)
```

Add name to each track

```{r}
xml_add_child(xml_children(doc), "name", c("track 1", "track 2"))
```

Add the `trkseg` nodes
```{r}
xml_add_child(xml_children(doc), "trkseg")
```

```{r}
doc
```

```{r}
trkseg_nodes <- xml_find_all(doc, "//trkseg")
for(k in 1:10) xml_add_child(trkseg_nodes, "trkpt")

```

```{r}
trkpt_nodes <- xml_find_all(trkseg_nodes, "//trkpt")
```

```{r}
xml_set_attr(trkpt_nodes, "lat", df$lat)
xml_set_attr(trkpt_nodes, "lon", df$lon)
xml_add_child(trkpt_nodes, "ele", df$ele)
xml_add_child(trkpt_nodes, "time", df$time)
```

```{r}
doc
```


```{r}
write_xml(doc, "example_1.gpx")
```


```{r}
xml2::xml_children(doc1)[[1]] %>% xml2::xml_children() %>% .[[1]]
```
Afegir un nou node `trk`

```{r}
xml2::xml_add_child(doc1, "trk")
```

```{r}
doc1
```
Afegir un enter (ele = 200) a dins de les coordenades del primer punt

```{r}
primer_punt_node <- 
  xml2::xml_children(doc1)[[1]] %>% xml2::xml_children() %>% .[[1]] %>% xml2::xml_children() %>% .[[1]]
```

```{r}
xml2::xml_add_child(primer_punt_node, "ele")
```

```{r}
ele_node <- primer_punt_node %>% xml2::xml_children() %>% .[[1]]
```

```{r}
xml2::xml_text(ele_node) <- 200
```

```{r}
xml2::xml_text(ele_node) <- '\"200\"'
```

```{r}
print(doc1)
```
## Create a track

```{r}
add_gpx <-  function(version = "1.1.0", creator = "R - pep"){
  xml_new_root("gpx",
  version = version,
  creator = creator,
  "xmlns:xsi"="http://www.w3.org/2001/XMLSchema-instance",
  xmlns="http://www.topografix.com/GPX/1/1",
  "xsi:schemaLocation"="http://www.topografix.com/GPX/1/1 http://www.topografix.com/GPX/1/1/gpx.xsd",
  "xmlns:gpxtpx"="http://www.garmin.com/xmlschemas/TrackPointExtension/v1")
}
```


```{r}
add_trk <- function(parent, df, name = NULL) {
  trk_node <- xml_add_child(parent, "trk")
  n_points <- nrow(df)
  if(!is.null(name)) xml_add_child(trk_node, "name", name)
  # add trk, name, trkseg, trkpt
  trkseg_node <- xml_add_child(trk_node, "trkseg")
  for(k in 1:n_points) {xml_add_child(trkseg_node, "trkpt")}
  trkpt_nodes <- xml_find_all(trkseg_node, ".//trkpt")
  # create attribute lon, lat
  xml_set_attr(trkpt_nodes, "lat", df$lat)
  xml_set_attr(trkpt_nodes, "lon", df$lon)
  # create nodes ele and time if they exist 
  if("ele" %in% names(df)) xml_add_child(trkpt_nodes, "ele", df$ele)
  if("time" %in% names(df)) xml_add_child(trkpt_nodes, "time", df$time)
  return(invisible(parent))
}
```

```{r}
doc1 <- add_gpx(creator = "pep")
add_trk(doc1, df, name = "track_example")
write_xml(doc1, file = "output_nou.gpx", options = c("format", "no_empty_tags"))
```
```{r}
doc2 <- add_gpx(creator = "pep")
add_trk(doc2, df[1:10, ], name = "track_1")
add_trk(doc2, df[11:20, ], name = "track_2")
write_xml(doc2, file = "output_doble.gpx", options = c("format", "no_empty_tags"))
```
## Third method using `sf_write`



## **XML** package


```{r}
doc2 <- XML::xmlParse('<gpx version="1.1" creator="R">
<trk>
<trkseg>
<trkpt lat="40.779" lon="-74.428" />
<trkpt lat="40.777" lon="-74.418" />
</trkseg>
 </trk>
 <trk>
 <trkseg>
<trkpt lat="50.779" lon="-64.428" />
<trkpt lat="50.777" lon="-64.418" />
</trkseg>
</trk>
</gpx>')
```

```{r}
print(doc2)
```



```{r}
XML::xmlSize(doc2)
```
```{r}
XML::xmlName(XML::xmlRoot(doc2))
```


```{r}
XML::xmlChildren(XML::xmlRoot(doc2)) 
```
```{r}
XML::xmlChildren(XML::xmlRoot(doc2)) %>% XML::xmlSize()
```

```{r}
XML::xmlChildren(XML::xmlRoot(doc2)) %>% .[[1]] %>% XML::xmlChildren() %>% .[[1]]
```

```{r}
first_point <- XML::getNodeSet(doc2, "//trkpt")
first_point_node = first_point[[1]]
```

```{r}
XML::xmlSize(first_point)
```
```{r}
first_point[[1]]
```

```{r}
first_point_node
```
```{r}
XML::newXMLNode("ele", '\"200\"' , parent = first_point_node)
```
```{r}
doc2
```



```{r}
XML::newXMLNode("ele", parent = first_point_node)
```
```{r}
XML::xmlValue(XML::xmlChildren(first_point_node)[[4]]) <- 500
```

```{r}
doc2
```


```{r}
XML::addChildren(first_point_node, "ele")
```
```{r}
ele_node_eq <- XML::xmlChildren(first_point_node)
```

```{r}
doc2
```

```{r}
XML::xmlValue(ele_node_eq[[1]])
```








