---
title: "Visualize GPX file information from Strava"
author: "Pep Porrà"
date: "2020-10-10"
output: 
  html_document:
    code_folding: "show"
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Goal

To visualize the information of GPX file exported from Strava (or other GPS software)

## Read file

```{r, message=F, warning=F}
library(tidyverse)
library(sf)
```

We read first the two pieces of information within the gpx file from Strava

```{r}
file_name = "data/strava/sant_ignasi_turo_de_monts_tibidabo_coll_de_la_vinassa_la_pinya_2_vallvidrera_santa_maria_de_jerusalem.gpx"
strava_sf_tracks = read_sf(file_name, layer = "tracks")
strava_sf_track_points = read_sf(file_name, layer = "track_points")
```
Many of the features are empty. We keep only those with information

```{r}
strava_sf_tracks = strava_sf_tracks %>%
  dplyr::select(name, type)
strava_sf_track_points = strava_sf_track_points %>%
  dplyr::select(track_seg_point_id, ele, time)
```

### Route length

From the tracks record, length can be computed using `st_length` function

```{r}
strava_sf_track_points
```

```{r}
n_points = nrow(strava_sf_track_points)
```


```{r}
st_length(strava_sf_tracks)
```
We can create a feature that gives the distance between a point and the precedent point. Function `st_distance` computes distances between object x and object y. We just shift object y to start at element 2 and object x to lack last point

```{r}
d_distances = st_distance(
  strava_sf_track_points[-n_points, ],
  strava_sf_track_points[-1, ], by_element = TRUE)
```

```{r}
sum(d_distances)
```
Note that strava_sf_track_points[-n_points, ] is an sf object with n_points-1 elements

```{r}
strava_sf_track_points[-n_points, ]
```

we can also compute the cumulative distances

```{r}
cum_distances = cumsum(d_distances)
```

and turn both into features in the sf object

```{r}
strava_sf_track_points$d_distances = c(units::set_units(0, m), d_distances)
strava_sf_track_points$cum_distances = c(units::set_units(0, m), cum_distances)
```

### Time invested 

We can computed the time passed to move form one point of the route to the next one

```{r}
time_segment = units::as_units(
  strava_sf_track_points$time[-1] - strava_sf_track_points$time[-n_points])
```

Total duration of the route reads

```{r}
total_time = sum(time_segment)
units::set_units(total_time, min) 
```

We can create two new features of the track_points object:

1. time per segment
2. cumulative time 

```{r}
strava_sf_track_points$time_segment = c(units::set_units(0, s), time_segment)
strava_sf_track_points$cum_time = c(units::set_units(0, s), cumsum(time_segment))
```

We can identify if there is any segment where time passed is zero

```{r}
zero_time_vector = units::set_units(rep(0.0, n_points - 1), s)
```

```{r}
sum(time_segment == zero_time_vector)
```
```{r}
which_zero = which(time_segment == zero_time_vector)
```

```{r}
d_distances[which_zero]
```
### Elevation gained

```{r}
elevation_segment = units::set_units(
  strava_sf_track_points$ele[-1] - strava_sf_track_points$ele[-n_points], m)
```

Total duration of the route reads

```{r}
total_elevation = sum(elevation_segment)
total_elevation 
```

We can create one new features of the track_points object:

1. elevation per segment
2. cumulative elevation 

```{r}
strava_sf_track_points$elevation_segment = c(units::set_units(0, m), elevation_segment)
strava_sf_track_points$cum_elevation = 
  c(units::set_units(0, m), cumsum(elevation_segment))
```

```{r}
plot(strava_sf_track_points$cum_distances,
  strava_sf_track_points$cum_elevation, col = "orange", pch = 16, cex = 0.5,
  xlab = "Distance [m]", ylab = "Elevation gain [m]", las = 1)
grid()
```





```{r}
st_dimension(strava_sf_tracks)
```
```{r}
str(strava_sf_tracks)
```
```{r}
strava_sf_tracks
```
```{r}
print(strava_sf_track_points, n = 3)
```
```{r}
coordenades_trkp = st_coordinates(strava_sf_track_points)
```

```{r}
coordenades_tr = st_coordinates(strava_sf_tracks)
```

```{r}
coordenades_tr[1:7, ]
```
```{r}
coordenades_trkp[1:7, ]
```
```{r}
st_geometry_type(strava_sf_tracks)
```
```{r}
st_geometry_type(strava_sf_track_points$geometry[[5]])
```
```{r}
class(strava_sf_tracks$geometry)
```
```{r}
class(strava_sf_track_points$geometry)
```
```{r}
linia = st_cast(strava_sf_tracks, to = 'LINESTRING')
```


```{r}
st_length(linia)
```

```{r}
resp = lapply(2:nrow(strava_sf_track_points), 
  function(k) st_linestring(coordenades_trkp[(k-1):k, ])
  )
```


## Simple example

```{r}
simple_line = st_linestring(matrix(c(0, 0, 1, 1), 2, 2, byrow = TRUE))
```

```{r}
plot(simple_line)
```
```{r}
st_crs(simple_line)
```
```{r}
simple_sfc = st_sfc(simple_line)
```


```{r}
st_crs(simple_sfc) = 4326
```

```{r}
st_length(simple_sfc)
```
```{r}
st_line_sample(
  st_transform(simple_sfc, 3857), 
  sample = c(0, 100000/as.numeric(st_length(simple_sfc))))
```
```{r}
st_cast(.Last.value, "LINESTRING")
```
```{r}
st_transform(.Last.value, 4326)
```
```{r}
linia_100km = .Last.value
```

```{r}
coordenades_punt =  st_coordinates(linia_100km)[2, 1:2]
point_100km = st_sfc(
  st_point(coordenades_punt),
  crs = 4326
)
```


```{r}
plot(simple_sfc)
plot(linia_100km, add = TRUE, col = "red", lwd = 2)
plot(point_100km, col = "red", pch = 16, add = TRUE, cex = 3)
```


```{r}
library(tmap)
```
```{r}
tm_shape(gps_track_points_sf) + tm_symbols(size = 0.1, col = "red") 
```

```{r}
tm_shape(gps_tracks_sf) + tm_lines(col = "orange", lwd = 2) +
tm_shape(gps_track_points_sf) + tm_symbols(size = 0.1, col = "red") 
```
```{r}
st_length(gps_tracks_sf)
```


```{r}
all_segments_distance = st_distance(gps_track_points_sf)
```

```{r}
m_distances  = matrix(c(1:112, 2:113), 112, 2)
```


```{r}
distances = c(units::set_units(0, "m"), all_segments_distance[m_distances])
sum(distances)
```
```{r}
st_line_sample(gps_tracks_sf, density = units::set_units(1, 1/km))
```

```{r}
st_geometry(gps_track_points_sf)[1:2]
```
```{r}
class(st_geometry(gps_track_points_sf)[1:2])
```

```{r}
st_crs(gps_track_points_sf)
```

```{r}
line_new = st_linestring(st_coordinates(st_geometry(gps_track_points_sf)[1:2]))
```

```{r}
line_new_sfc = st_sfc(line_new, crs = 4326)
```

```{r}
line_new_sfc
```


```{r}
st_line_sample(line_new, sample = c(0, 0.25, 0.5, 0.75, 1))
```
```{r}
st_line_sample(st_transform(line_new_sfc, 3857), density = units::set_units(1, 1/m))
```

