---
title: "Simple geometries"
author: "Pep Porrà"
date: "2020-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

To give examples of using simple geometry objects (geospatial vectors):

* sfg (Simple feature geometries)
* sfc (simple feature geometry list-columns)
* sf (Simple features)

## **sf** Package

```{r}
library(tidyverse)
library(sf)
library(ggplot2)
```

Main Reference to understand the concepts used in these examples is the first vignette of the **sf** package by Edzer Pebesma [link](https://cran.r-project.org/web/packages/sf/vignettes/sf1.html) 

## Structures

* First: Simple geometries POINTS, LINESTRINGS, POLYGONS, ... (sfg objects, simple feature geometry)
* Second: Simple geometry columns (sfc)
* Third: Simple features (sf objects)

### **s**imple **f**eature **g**eometry

#### POINT / MULTIPOINT

```{r}
p1 = st_point(c(1, 2))
p1
```

```{r}
p2 = st_point(c(1, 2, 3, 4))
p2
```
```{r}
p2 = st_point(5:6)
```

```{r}
class(p2)
```
```{r}
methods(class = "POINT")
```

```{r}
p3 = st_point(rep(0,2))
```

```{r}
n_points = c(p1, p2, p3)
```

Several points as sfg are a MULTIPOINT

```{r}
class(n_points)
```


```{r}
plot(n_points, axes = T, las = 1)
grid()
plot(n_points, pch = 16, col = "red", add = TRUE)
text(st_coordinates(p1) + 0.5 * c(1, 0), "p1")
text(st_coordinates(p2) + 0.5 * c(1, 0), "p2")
text(st_coordinates(p3) + 0.5 * c(1, 0), "p3")
```

```{r}
ggplot(n_points) + geom_sf(col = "red")
```


Other ways to build a MULTIPOINT

```{r}
st_multipoint(c(p1, p2))
```

```{r}
st_multipoint(st_coordinates(c(p1, p2))[, 1:2])
```
```{r}
st_multipoint(rbind(st_coordinates(p1), st_coordinates(p2)))
```

```{r}
st_multipoint(rbind(p1, p2))
```
```{r}
st_multipoint(as.matrix(c(p1, p2)))
```

```{r}
methods(class = "sfg")
```

```{r}
set.seed(1023)
mp1 = st_multipoint(matrix((1:8) + runif(8), 4, 2, byrow = T))
mp1

```

#### LINESTRING / MULTILINESTRING

```{r}
(
  l1 = st_linestring(c(p1, p2))
)
```
```{r}
(
  l2 = st_linestring(c(p1, p3))
)
```
```{r}
(
  l3 = st_linestring(mp1)
)
```
```{r}
ml1 = st_multilinestring(list(l1, l2, l3))
```

```{r}
plot(ml1, axes = TRUE)
plot(st_cast(ml1, "MULTIPOINT"), pch = 16, col = "red", add = TRUE)
```

#### POLYGON / MULTIPOLYGON

Create a function that distributes n equidistant points in a circle.

```{r}
n_pts_unit <- function(n, polygon = FALSE, r = 1, start = pi / 2){
  
  angles = 2 * pi * ((1:n) - 1) / n + start
  x = round(r * cos(angles), 8)
  y = round(r * sin(angles), 8)
  mp = matrix(c(x, y), n, 2)
  if (polygon) mp = rbind(mp, mp[1, ])
  result = st_multipoint(mp)
  return(result)
} 
```

```{r}
points_p1 = n_pts_unit(11)
points_p1 = rbind(points_p1, points_p1[1, ])
(
  pp1 = st_polygon(list(points_p1))
)
```
```{r}
plot(pp1, axes = TRUE)
plot(st_cast(pp1, "MULTIPOINT"), pch = 16, col = "red", add = TRUE)
```


```{r}
points_p2 = n_pts_unit(5)
points_p2 = rbind(points_p2, points_p2[1, ])

(
  pp2 = st_polygon(list(points_p2, 0.8 * points_p2))
)
```
```{r}
plot(pp2, col = 'grey')
```
```{r}
plot(
  st_multipolygon(
    list(
      list(n_pts_unit(4, polygon = T, 2)),
      list(n_pts_unit(6, polygon = T, 2)),
      list(n_pts_unit(9, polygon = T, 2))
    )
  )
)
```

```{r}
old_par = par(mar = rep(0,4))
n = 11
plot(
  st_multipolygon(
    list(
      lapply( ((1:n) - 1) * 2 * pi / n , 
        function(alpha) n_pts_unit(9, polygon = T, 
          r = 2 - 2 * alpha / (2 * pi) , start = alpha))
    )
  )
)
par(old_par)
```


## **s**imple **f**eature **c**olumn

sfg objects can be turn into sfc objects that have a coordinate reference system (CRS attribute). All the geometry information of sf objects is contained in an object of type sfc

```{r}
(sc_mp1 =st_sfc(mp1))
```

```{r}
st_coordinates(sc_mp1)
```



We can use `st_cast` function to simplify a geometry. For instance, obtain teh points in a multipoint object as an object with n points geometry

```{r}
(
  sc_n_p1 = st_cast(sc_mp1, "POINT")
)
```
All sfg elements of a sfc are of equal type and each of them correspond to an element of a list. For instance, sc_n_p1 has 4 features. Therefore, sc_n_p1[[2]] gives us the sfg of the second element and it is of class "sfg"


```{r}
sc_n_p1[[2]]
```
```{r}
sc_n_p1[[2]] %>% class()
```



```{r}
st_coordinates(sc_n_p1)
```
We can also build linestrings for multipoints or points

```{r}
st_cast(sc_mp1, "LINESTRING")
```
From points to multipoints, use `st_union`

```{r}
st_union(
  st_cast(sc_mp1, "POINT")
)
```

```{r}
sc_mp1
```


```{r}
st_cast(st_union(sc_n_p1[1], sc_n_p1[2]), "LINESTRING")
```
```{r}
st_linestring(c(sc_n_p1[[1]], sc_n_p1[[2]]))
```

```{r}
(
  segments_n_p1 = 
  lapply(1:(length(sc_n_p1) - 1), 
    function(k) st_linestring(c(sc_n_p1[[k]], sc_n_p1[[k + 1]])))
)
```
```{r}
segments_n_p1[[1]]
```
```{r}
st_sfc(segments_n_p1)
```
A function to transform a sfc with n POINTs to a sfc with the n-1 segments created sequentially for each pair of segments reads

```{r}
points_to_segments = function(sfc){
  n_segments = length(sfc) - 1
  crs = st_crs(sfc)
  
  result = st_sfc(
    lapply(1:n_segments, 
      function(k) st_linestring(c(sfc[[k]], sfc[[k + 1]]))),
    crs = crs
  )
  return(result)
}

```

```{r}
points_to_segments(sc_n_p1)
```
This function is useful when working with routes in gpx files.


```{r}
(
  sc_lp1 = st_cast(sc_mp1, "LINESTRING")
)
```
```{r}
st_length(sc_lp1)
```
```{r}
st_cast(sc_lp1, 'MULTIPOINT')
```


```{r}
plot(sc_lp1, axes = TRUE)
plot(st_cast(sc_lp1, 'POINT'),
  pch = 16, col = "red", add = TRUE)
```


```{r}
st_segmentize(sc_lp1, 1)
```
```{r}
plot(sc_lp1, axes = TRUE)
plot(st_cast(sc_lp1, 'POINT'),
  pch = 16, col = "red", add = TRUE)
plot(st_cast(
    st_segmentize(sc_lp1, 2.5), 'POINT'),
  pch = 1, col = "green", add = TRUE, cex = 2)
```

Function to find points at given units length

```{r}
total_distance = st_length(sc_lp1)
unit_length = 4
n_exact_distance = floor(total_distance / unit_length)
st_line_sample(sc_lp1, 
  sample = c(0, unit_length * (1:n_exact_distance), total_distance)/total_distance)
```
```{r}
plot(sc_lp1, axes = TRUE)
plot(st_cast(sc_lp1, 'POINT'),
  pch = 16, col = "red", add = TRUE)
plot(
  st_cast(
    st_line_sample(sc_lp1, 
      sample = c(0, unit_length * (1:n_exact_distance), total_distance)/total_distance),
    'POINT'),
  pch = 16, col = "green", add = TRUE)
```

## **sf** objects

On top of the geometric information, we add columns that have information for each featurein the sfc part. For instance, in order to convert the sfc object sc_n_p1 into an sf object we define a column with the id of each feature.

```{r}
(
  sf_n_p1 = st_sf(id = 1:length(sc_n_p1), sc_n_p1)
)
```

```{r}
plot(sf_n_p1, axes = T)
```

```{r}
ggplot(sf_n_p1) + geom_sf(aes(color = as.factor(id)))
```

