---
title: "temp"
author: "Pep Porrà"
date: "20/4/2021"
output: html_document
---



```{r}
sfc_tracks_min <- st_geometry(sf_tracks_min)

```

```{r}
tracks_coord <- st_coordinates(sfc_tracks_min)
```


```{r}
class(tracks_coord)
```
```{r}
str(tracks_coord)
```

```{r}
str(tracks_coord)
```

```{r}
sf_tracks_min_1 <- st_sf(
  st_sfc(
    st_linestring(
      st_multipoint(
        x=tracks_coord[-13, 1:2]))
    , crs = st_crs(sf_tracks_min))
)
```
```{r}
class(tracks_1_sfc)
```

```{r}
head(tracks_coord_sfg)
```



```{r}
st_multipoint(matrix(1:4, 2, byrow = TRUE))
```
```{r}
m <- matrix(1:10, 5, byrow = TRUE)
st_linestring(m)
```
```{r}
str(m)
```



```{r}
tracks_coord_1 <- st_multipoint(tracks_coord)

```

```{r}
tracks_1_sfg <- st_linestring(tracks_coord_1)
tracks_1_sfc <-  st_sfc(tracks_1_sfg, crs = st_crs(sfc_tracks_min))
```



```{r}
coord_track <-  st_coordinates(sf_tracks)
```

```{r}
coord_points_track <-  st_coordinates(sf_track_points)
```




```{r}
st_crs(sf_track_points)
```

```{r}
st_crs(sf_tracks)
```




We delete point 13:

```{r}
sh_sf_modified <- sh_sf_track_points[-13, ]
```


```{r}
plot(sh_sf_modified[1:30, "ele"], type = "l", col = "red", lwd = 2)
```


```{r}
st_write(sh_sf_modified[, c("ele")], "data/track_modified_2.gpx", driver = "GPX")
```
### To be done: Delete a point from the track (multistring)
### Read the multistring element of an sf object



```{r}
str(sh_sf_track_points)
```

```{r}
st_bbox(sh_sf_track_points)
st_bbox(sh_sf_track_points[1:30, ])
```

### Route length

From the tracks record, length can be computed using `st_length` function

```{r}
sh_sf_track_points
```

```{r}
n_points = nrow(sh_sf_track_points)
```


```{r}
st_length(sh_sf_tracks)
```
We can create a feature that gives the distance between a point and the precedent point. Function `st_distance` computes distances between object x and object y. We just shift object y to start at element 2 and object x to lack last point

```{r}
d_distances = st_distance(
  sh_sf_track_points[-n_points, ],
  sh_sf_track_points[-1, ], by_element = TRUE)
```


```{r}
sum(d_distances)
```
Note that sh_sf_track_points[-n_points, ] is an sf object with n_points-1 elements

```{r}
sh_sf_track_points[-n_points, ]
```

we can also compute the cumulative distances

```{r}
cum_distances = cumsum(d_distances)
```

and turn both into features in the sf object

```{r}
sh_sf_track_points$d_distances = c(units::set_units(0, m), d_distances)
sh_sf_track_points$cum_distances = c(units::set_units(0, m), cum_distances)
```

```{r}
sh_sf_track_points %>% 
  ggplot(aes(x=time, y=as.numeric(d_distances))) + 
  geom_point()
```


### Time invested 

We can computed the time passed to move form one point of the route to the next one

```{r}
time_segment = units::as_units(
  sh_sf_track_points$time[-1] - sh_sf_track_points$time[-n_points])
```

Total duration of the route reads

```{r}
total_time = sum(time_segment)
units::set_units(total_time, min) 
```

We can create two new features of the track_points object:

1. time per segment
2. cumulative time 

```{r}
sh_sf_track_points$time_segment = c(units::set_units(0, s), time_segment)
sh_sf_track_points$cum_time = c(units::set_units(0, s), cumsum(time_segment))
```

```{r}
time_segment[time_segment == units::set_units(0, s)]
```
```{r}
table(time_segment)
```





```{r}
sh_sf_track_points %>% 
  ggplot(aes(x=as.numeric(time_segment), y=as.numeric(d_distances))) + 
  geom_point()
```
```{r}
sh_sf_track_points %>% 
  ggplot(aes(x=time, y=1/(as.numeric(d_distances/time_segment) * 60 / 1000))) + 
  geom_line() +
  coord_cartesian(ylim = c(0,25))
```
```{r}
head(sort(as.numeric(d_distances/time_segment), decreasing = T), n = 20)
```
```{r}
sh_sf_track_points %>%
  arrange(desc(d_distances)) %>% 
  select(d_distances, cum_distances, track_seg_point_id)
```

```{r}
sh_sf_track_points %>%
  arrange(d_distances) %>% 
  select(d_distances, cum_distances, time_segment, cum_time, track_seg_point_id)
```



```{r}
sh_sf_track_points %>% 
  ggplot(aes(x=time, y=as.numeric(d_distances/time_segment))) + 
  geom_line() +
  coord_cartesian(ylim = c(0,25))
```

```{r}
plot(1:1049, time_segment)
```

```{r}
plot(1:1049, d_distances, ylim = c(10,40))
```

We can identify if there is any segment where time passed is zero

```{r}
zero_time_vector = units::set_units(rep(0.0, n_points - 1), s)
```

```{r}
sum(time_segment == zero_time_vector)
```
```{r}
which_zero = which(time_segment == zero_time_vector)
```

```{r}
d_distances[which_zero]
```

### Elevation gained

```{r}
elev <- sh_sf_track_points$ele
elev <- elev[!is.na(elev)]
elev_1 <- elev[-1]
elev_0 <- elev_0[-n_points]
elevation_segment  <-  units::set_units(elev_1 - elev_0, m)
```

Total duration of the route reads

```{r}
total_elevation = sum(elevation_segment)
total_elevation 
```
```{r}
plot(1:329, elev)
```




We can create one new features of the track_points object:

1. elevation per segment
2. cumulative elevation 

```{r}
strava_sf_track_points$elevation_segment = c(units::set_units(0, m), elevation_segment)
strava_sf_track_points$cum_elevation = 
  c(units::set_units(0, m), cumsum(elevation_segment))
```

```{r}
plot(strava_sf_track_points$cum_distances,
  strava_sf_track_points$cum_elevation, col = "orange", pch = 16, cex = 0.5,
  xlab = "Distance [m]", ylab = "Elevation gain [m]", las = 1)
grid()
```





```{r}
st_dimension(strava_sf_tracks)
```
```{r}
str(strava_sf_tracks)
```
```{r}
strava_sf_tracks
```
```{r}
print(strava_sf_track_points, n = 3)
```
```{r}
coordenades_trkp = st_coordinates(strava_sf_track_points)
```

```{r}
coordenades_tr = st_coordinates(strava_sf_tracks)
```

```{r}
coordenades_tr[1:7, ]
```
```{r}
coordenades_trkp[1:7, ]
```
```{r}
st_geometry_type(strava_sf_tracks)
```
```{r}
st_geometry_type(strava_sf_track_points$geometry[[5]])
```
```{r}
class(strava_sf_tracks$geometry)
```
```{r}
class(strava_sf_track_points$geometry)
```
```{r}
linia = st_cast(strava_sf_tracks, to = 'LINESTRING')
```


```{r}
st_length(linia)
```

```{r}
resp = lapply(2:nrow(strava_sf_track_points), 
  function(k) st_linestring(coordenades_trkp[(k-1):k, ])
  )
```


## Simple example

```{r}
simple_line = st_linestring(matrix(c(0, 0, 1, 1), 2, 2, byrow = TRUE))
```

```{r}
plot(simple_line)
```
```{r}
st_crs(simple_line)
```
```{r}
simple_sfc = st_sfc(simple_line)
```


```{r}
st_crs(simple_sfc) = 4326
```

```{r}
st_length(simple_sfc)
```
```{r}
st_line_sample(
  st_transform(simple_sfc, 3857), 
  sample = c(0, 100000/as.numeric(st_length(simple_sfc))))
```
```{r}
st_cast(.Last.value, "LINESTRING")
```
```{r}
st_transform(.Last.value, 4326)
```
```{r}
linia_100km = .Last.value
```

```{r}
coordenades_punt =  st_coordinates(linia_100km)[2, 1:2]
point_100km = st_sfc(
  st_point(coordenades_punt),
  crs = 4326
)
```


```{r}
plot(simple_sfc)
plot(linia_100km, add = TRUE, col = "red", lwd = 2)
plot(point_100km, col = "red", pch = 16, add = TRUE, cex = 3)
```


```{r}
library(tmap)
```
```{r}
tm_shape(gps_track_points_sf) + tm_symbols(size = 0.1, col = "red") 
```

```{r}
tm_shape(gps_tracks_sf) + tm_lines(col = "orange", lwd = 2) +
tm_shape(gps_track_points_sf) + tm_symbols(size = 0.1, col = "red") 
```
```{r}
st_length(gps_tracks_sf)
```


```{r}
all_segments_distance = st_distance(gps_track_points_sf)
```

```{r}
m_distances  = matrix(c(1:112, 2:113), 112, 2)
```


```{r}
distances = c(units::set_units(0, "m"), all_segments_distance[m_distances])
sum(distances)
```
```{r}
st_line_sample(gps_tracks_sf, density = units::set_units(1, 1/km))
```

```{r}
st_geometry(gps_track_points_sf)[1:2]
```
```{r}
class(st_geometry(gps_track_points_sf)[1:2])
```

```{r}
st_crs(gps_track_points_sf)
```

```{r}
line_new = st_linestring(st_coordinates(st_geometry(gps_track_points_sf)[1:2]))
```

```{r}
line_new_sfc = st_sfc(line_new, crs = 4326)
```

```{r}
line_new_sfc
```


```{r}
st_line_sample(line_new, sample = c(0, 0.25, 0.5, 0.75, 1))
```
```{r}
st_line_sample(st_transform(line_new_sfc, 3857), density = units::set_units(1, 1/m))
```

